Lens Modeling in \sw

### 2013-12-12 text/tex/abstract.tex
more edits

\begin{abstract}
In Spacewarps **, volunteers are invited to search sky surveys for lens
candidates.  Here we report on a system that allows experienced
volunteers to model lens candidates.  A sample of 29 simulated lenses
were modeled, each multiple times.  The quality of the models was
then examined in different ways.  In particular, (i)~the image
parities and time orderings were identified correctly in most cases,
but not always; (ii)~the mean convergence (equivalent to the enclosed
mass) was quite well constrained in the image region; (iii)~the
performance of experienced volunteers was comparable to professionals.
\end{abstract}

>>>>

\begin{abstract}
In Spacewarps **, volunteers are invited to search sky surveys for
lens candidates.  Here we report on a system that allows experienced
volunteers to model lens candidates.  A sample of 29 simulated lenses
were modeled, each multiple times.  The quality of the models was then
examined in different ways.  In particular: (i)~the image parities and
time orderings were identified correctly in most cases, but not
always; (ii)~the mean convergence (equivalent to the enclosed mass)
was quite well constrained in the image region; (iii)~the performance
of experienced volunteers was comparable to professionals.
\end{abstract}

### 2013-12-12 text/tex/outlook.tex
more edits

One, model the source as well.

Another, lenses with more than one maximum can be allowed for.  See
Figure 5c in \citep{2001ApJ...557..594R} and Figure 4b in
\cite{2003ApJ...590...39K}.

Other modelling approaches also possible.

As seen in \secref{tests.t1}, volunteers fail mostly in two situations:
when to identify an arc like structure while placing the points
and in 5 image configurations, that are quite symmetric, then to identify the correct ordering of the points.

The first can be improved by additional images from different filters (?)
Will be implemented with a next major \spl upgrade.

The second can be made less by better training (introduce the ruler), may be more filters help too?

We hope that those problems are also taken care of when staring collaborative modeling. (Already in progress)
Basics are available ('revise functionality') but better community tools would help (having overview, comparison between different models etc...)

>>>>

\begin{itemize}
\item Currently, \spl does not attempt to model the source shape; the
  user identifies the brightest points on the image, and these are
  taken as the images of a point-like source, whose positions must be
  reproduced exactly.  For generating a synthetic image, a conical
  source profile is assumed.  Fitting for the source profile to
  optimize resemblance to the observed lensed image after the lens
  model has been generated, is algorithmically straightforward and
  planned to be implemented.  This would alleviate another problem
  with \spl, which is that there is no numerical figure of merit, and
  assessment of a model is a judgement call based on the synthetic
  image, and on whether the mass distribution and the arrival-time
  surface show suspect features.
\item Another current limitation in \spl is that the lens is assumed
  to be dominated by one galaxy, which puts most galaxy-group lenses
  beyond the reach of the modeller.  Since complicated group lenses
  are some of the most interesting candidates present, removing this
  limitation is most desirable.  From the users' point of view, it
  would mean that spaghetti contours with more than one maximum can be
  allowed.  For examples, see Figure 5c in \citep{2001ApJ...557..594R}
  and Figure 4b in \cite{2003ApJ...590...39K}.
\item Currently, a single false-colour composite is used as the data.
  An option could be added to use all available filters, individually
  or in cmbination, as the user wishes.
\end{itemize}

The third desirable avenue of improvement is to facilitate
collaborative work.

\begin{itemize}
\item As mentioned above, the option of revising an already-archived
  model is already available.  Desired now are tools for comparing
  different models of a given system, both visually and through
  different statistical measures.  As evidenced by a current
  collaborative modelling effort, a particularly interesting candidate
  can lead to an extended discussion and dozens of models.
\item Better tutorial materials are also needed, and this would
  address some of the problem areas found in the modelling challenge.
  For example, we saw in \secref{tests.t1} that volunteers are most
  prone to making errors in two situations: when to identify an arc-
  like structure while placing the points, and in identifying the
  correct ordering of the points in nearly-symmetric configurations.
\end{itemize}
